---
title: "TRSL: regression against translation rates"
author: "Martin Seeger"
date: "18. April 2016"
output: html_document
---

```{r, echo=FALSE}
# prepare data frame
#setwd("~/git/TRSL/parameters")
# load data from csv
data = read.csv("regression_data.csv", sep=",")
# rename columns
colnames(data) = c("X", "gene", "translation.rate.exp", "translation.rate.sim", "transcript.abundance", "initiation.rate", "ORF.length", "CAI")
```
## Unnormalized models

### Simulated translation rates

```{r}
model.unnorm.sim = lm(data$translation.rate.sim ~ data$initiation.rate + data$CAI)
summary(model.unnorm.sim)
```

### Experimental translation rates

Data from [Arava et al.](http://www.pnas.org/content/100/7/3889.full#ref-list-1):

```{r}
model.unnorm.exp = lm(data$translation.rate.exp ~ data$initiation.rate + data$CAI)
summary(model.unnorm.exp)
```

The coefficients are not comparable because the factors have different orders of magnitude.

Main problem: poor fit.

We scale (standard normalize) the variables:
```{r}
IR_norm = scale(data$initiation.rate)
CAI_norm = scale(data$CAI)
```

## Normalized models

### Simulated translation rates

```{r}
model.sim = lm(data$translation.rate.sim ~ IR_norm + CAI_norm)
summary(model.sim)
```

### Experimental translation rates

```{r}
model.exp = lm(data$translation.rate.exp ~ IR_norm + CAI_norm)
summary(model.exp)
```

Unsurprisingly, same quality ($p$, $R^2$) as for the unnormalized models.

Create matrix of values, only complete cases. We include the dependent variables for later plotting.

```{r}
factors_norm = cbind(IR_norm, CAI_norm, data$translation.rate.sim, data$translation.rate.exp)[complete.cases(cbind(IR_norm, CAI_norm)),]
# add vector of ones for score calculation
factors_norm = cbind(rep(1, nrow(factors_norm)), factors_norm)
colnames(factors_norm) = c('const', 'IR_norm', 'CAI_norm', 'translation.rate.sim', 'translation.rate.exp')
```

Define a score $s = t + b_1 \cdot x_1 + b_2 \cdot x_2$:

```{r}
scores.sim = factors_norm[, c('const', 'IR_norm', 'CAI_norm')] %*% model.sim$coefficients
plot.new()
curve(x^1, from=0, to=0.5, col="red", xlab="score", ylab="translation rate (sim.)")
points(scores.sim, factors_norm[, "translation.rate.sim"], ylim=c(0,50), main="", pch=1, col='blue')
```

```{r}
scores.exp = factors_norm[, c('const', 'IR_norm', 'CAI_norm')] %*% model.exp$coefficients
plot.new()
curve(x^1, from=0, to=0.5, col="red", xlab="score", ylab="translation rate (exp.)")
points(scores.exp, factors_norm[, "translation.rate.exp"], ylim=c(0,50), main="", pch=1, col='blue')
```

This is obviously not a good model.


